\chapter{Installation}\label{sec:install}
\index{installation}

\section{Requirements}\label{sec:install.requirements}
\OOMMF\ software is written in \Cplusplus\ and \Tcl.  It uses the \Tcl-based
\Tk\ Windowing Toolkit to create graphical user interfaces that are
portable to many varieties of \Unix, \Windows, and \MacOSX.

\Tcl\ and \Tk\ must be installed before installing \OOMMF.
\Tcl\ and \Tk\ are available for free from the
\htmladdnormallinkfoot{Tcl Developer Xchange}{https://www.tcl-lang.org/}.
We recommend the latest stable
versions of \Tcl\ and \Tk\ concurrent with this release of \OOMMF.
\OOMMF\ requires\index{requirement!Tcl/Tk}
at least version 8.5 of \Tcl\ and \Tk.  \OOMMF\ software
does not support any alpha or beta versions of \Tcl/\Tk, and
each release of \OOMMF\ may not work with later releases of
\Tcl/\Tk.  Check the release dates of both \OOMMF\ and
\Tcl/\Tk\ to ensure compatibility.

A \Tcl/\Tk\ installation includes two shell programs.  The names of
these programs may vary depending on the \Tcl/\Tk\ version and the
type of platform.  The first shell program contains an interpreter
for the base \Tcl\ language.  In the \OOMMF\ documentation we refer
to this program as \fn{tclsh}\index{application!tclsh}.
The second shell program contains
an interpreter for the base \Tcl\ language extended by additional
\Tcl\ commands supplied by the \Tk\ toolkit.  In the
\OOMMF\ documentation we refer to this program as
\fn{wish}\index{application!wish}.
Consult your \Tcl/\Tk\ documentation to determine
the actual names of these programs on your platform (for example,
\fn{tclsh86.exe} or \fn{wish8.6}).

\OOMMF\ applications communicate via TCP/IP\index{TCP/IP} network sockets.
This means that \OOMMF\ requires\index{requirement!TCP/IP}
support for networking, even
on a stand-alone machine.  At a minimum, \OOMMF\ must be able to
access the loopback interface so that the host can talk to
itself using TCP/IP.

\index{application!Xvfb|(}
\OOMMF\ applications that use \Tk\ require a windowing system and
a valid display.  On Unix systems, this means that an X server must
be running.  If you need to run \OOMMF\ applications on a Unix system
without display hardware or software, you may need to start the
application with command line option
\hyperrefhtml{\texttt{\textbf{-tk 0}}}{\textsf{\textbf{-tk 0}} (see Ch.~}{)}{sec:cll}
or use the
\htmladdnormallinkfoot{Xvfb}{https://www.x.org/archive/X11R7.6/doc/man/man1/Xvfb.1.xhtml}
virtual frame buffer.
\index{application!Xvfb|)}

%\index{requirement!disk~space|(}
%The \OOMMF\ source distribution unpacks into a directory tree containing
%about 1500 files and directories, occupying approximately 80 MB of
%storage.  The amount of disk space needed for compiling and linking
%varies greatly between platforms; allow up to an additional 60 MB for
%the build.  Removing intermediate object modules (cf.\ the \cd{pimake}
%``objclean'' target, in \html{the} \hyperrefhtml{Reducing Disk Space
%Usage}{Reducing Disk Space Usage, Sec.~}{,
%below}{sec:install.reducedisk} \html{section}) reduces the final space
%requirement for source + binary executables to between 80 MB and 90 MB.
%The \OOMMF\ distribution containing \Windows\ executables unpacks into a
%directory tree occupying about 85 MB of storage.
%{\bf Note:} On a
%non-compressed {\tt FAT16} file system on a large disk, \OOMMF\ may take
%up much more disk space.  This is because on such systems, the minimum
%size of any file is large, as much as 32 KB.  Since this is much larger
%than many files in the \OOMMF\ distribution require, a great deal of
%disk space is wasted.
%\index{requirement!disk~space|)}

\index{compilers}\index{platforms}
To build \OOMMF\ software from source code, you will
need\index{requirement!C++~compiler} a \Cplusplus\ compiler
that implements the features of the C++11 standard.
%capable of handling
%C++ templates, C++ exceptions, and (for the \OOMMF\ eXtensible Solver)
%the C++ Standard Template Library.
You will need other software
development utilities for your platform as well.
We do development and
test builds on the following platforms, although porting to others
should not be too difficult:

% Note 1: Don't use transparent images, because mmHelp renders them
%   rather slowly.
% Note 2: makeimage (maybe +tabular?) breaks on newer systems, with
%   latex2html v1.71 and v1.68.  OTOH, v1.68 works on older systems,
%   so it is not clear what the problem is.  The error message is:
%          panic: end_shift at /usr/local/bin/latex2html line 11720.
%   This is with Perl v5.8.0.  So, we have to render the table
%   directly.  This is actually preferred with real HTML browsers,
%   but as of this writing (Dec-2004) mmHelp doesn't do tables.
% Note 3: Update, Mar-2021: mmHelp does tables now, so one can build
%   with or without the \begin{makeimage}/\end{makeimage}
%   environment. The table renders a bit cleaner with makeimage, but
%   some versions of latex2html don't seem to handle the makeimage
%   environment. Specifically, the version reporting to be "Version 2012
%   (1.2)" doesn't work. OTOH, on some machines we have installed
%   "Version 2018" which handles the makeimage enviroment fine.
\begin{center}
%\html{\begin{makeimage}\htmlimage{no_transparent}}
\begin{tabular}{|l|l|}\hline
Platform & Compilers \\ \hline
% AIX & VisualAge C++ (xlC), Gnu g++ \\
%Alpha/Linux & Compaq C++, Gnu g++ \\
%Alpha/Tru64 & Compaq C++ (cxx) \\
%Alpha/Windows NT & Microsoft Visual C++ \\
%HP-UX & aCC \\
Windows &
Microsoft Visual C++, MinGW g++, Cygwin g++
%Microsoft Visual C++, Borland C++, \\
% & Digital Mars dmc, MinGW g++, Cygwin
%, Intel C++
\\
Linux/x86 & Gnu g++, Intel C++
% , Portland Group pgCC
\\
%Linux/Itanium & Intel C++, Gnu g++ \\
\MacOSX & Clang C++, Gnu g++\\
%MIPS/IRIX 6 (SGI) & MIPSpro C++ \\
%SPARC/Solaris & Sun Workshop C++ \\
\hline
\end{tabular}
%\html{\end{makeimage}}
\end{center}

\section{Basic Installation}\label{sec:install.basic}
% Note: LaTeXML uses \label value to set file names.

Follow the instructions in the following sections, in order,
to prepare \OOMMF\ software for use on your computer.

\subsection{Download}\label{sec:install.download}
% Note: LaTeXML uses \label value to set file names.
\index{download}

The latest release of the \OOMMF\ software may be retrieved from the
\htmladdnormallinkfoot{\OOMMF\ download
page}{https://math.nist.gov/oommf/software.html}.  Each release is
available in two formats.  The first format is a gzipped tar file
containing an archive of all the \OOMMF\ source code.  The second format
is a \fn{.zip} compressed archive containing source code and
pre-compiled executables for \Windows.  Each \Windows\ binary
distribution is compatible with only a particular sequence of releases
of \Tcl/\Tk.  For example, a Windows binary release for \Tcl/\Tk\ 8.6.x
is compatible with \Tcl/\Tk\ 8.6.0, 8.6.1, \ldots.
%Other release
%formats, e.g., pre-compiled executables for Microsoft
%\Windows~NT running on a Compaq Alpha Systems RISC-based microprocessor
%system, and/or compatible with older versions of \Tcl/\Tk, may be made
%available upon request.

For the first format, unpack the distribution archive using gunzip and
tar:
\begin{verbatim}
gunzip -c oommf20a0.tar.gz | tar xvf -
\end{verbatim}

For the other format(s), you will need a utility program to unpack the
\fn{.zip} archive.
%This program must preserve the directory structure
%of the files in the archive, and it must be able to generate files with
%names not limited to the old MSDOS 8.3 format.  Some very old
%versions of the pkzip utility do not have these properties.
One utility program which is known to be suitable is
\htmladdnormallinkfoot{UnZip}{http://www.info-zip.org/pub/infozip/UnZip.html}.

Using your utility, unpack the \fn{.zip} archive, e.g.
\begin{verbatim}
unzip oommf20a0_86.zip
\end{verbatim}

For either distribution format, the unpacking sequence creates a
subdirectory \fn{oommf} which contains all the files and directories
of the \OOMMF\ distribution.  If a subdirectory named \fn{oommf}
already existed (say, from an earlier \OOMMF\ release), then
files in the new distribution overwrite those of the same name already
on the disk.  Some care may be needed in that circumstance to be
sure that the resulting mix of files from an old and a new
\OOMMF\ distribution combine to create a working set of files.

\subsection{Effects of the Installed \Tcl/\Tk}\label{sec:install.tcltk}
% Note: LaTeXML uses \label value to set file names.
\OOMMF\ interacts with your \Tcl/\Tk\ installation in several ways.  One
important restriction is that the major+minor release number of
\Tcl/\Tk\ must match the major+minor release number of the \Tcl/\Tk\
that \OOMMF\ was built against.  For example, if \OOMMF\ was built using
\Tcl/\Tk\ 8.5.18, then the resulting executables can run with any past or
future releases of \Tcl/\Tk\ from the 8.5.* series, but they won't run
(for example) with \Tcl/\Tk\ 8.4.20 or 8.6.4.

Another restriction is that the width of memory addresses in \Tcl/\Tk\
and \OOMMF\ must match.  Most general-purpose operating systems today
use primarily 64-bit memory addresses, but for backwards compatibility
can also run programs using 32-bit memory addresses.  However, a 64-bit
executable cannot link against a 32-bit library, or vice versa.
Therefore, if you have a 64-bit \Tcl/\Tk\ installed, then you will need
a 64-bit \OOMMF, and likewise a 32-bit \Tcl/\Tk\ needs a 32-bit \OOMMF.

Another restriction is that while \OOMMF\ can be built to run in
parallel across multiple cpu cores on a shared memory machine using
threads, to do this requires that the installed \Tcl/Tk\ be
thread-enabled.  Typical \Tcl/Tk\ installs on \Windows\ and \MacOSX\ are
thread-enabled.  \Tcl/Tk\ installs on recent releases of \Unix\ also
tend to be thread-enabled, but some older versions have non-threaded
\Tcl/Tk\ installs.  If your system \Tcl/\Tk\ install is non-threaded,
then you can either build a non-threaded version of \OOMMF, or else you
can make an additional, threaded \Tcl/\Tk\ install, for example under
your home directory or \fn{/usr/local}.  Be aware that if you have
multiple \Tcl/\Tk\ installations on your system then you need to be
careful to use the proper \fn{tclsh} whenever you build or launch
\OOMMF.

If you download \OOMMF\ with pre-built binaries, then it is imperative
that you select the download that matches the major+minor release
number and memory address width of the \Tcl/\Tk\ you want to run
\OOMMF\ with.  On the other hand, if you build \OOMMF\ from source,
then the \fn{tclsh} you use to run the build process is inspected to
determine relevant information about the local \Tcl/\Tk\ environment.
%If you are using one of the \OOMMF\ tested
%\hyperrefhtml{platform+compiler combinations}{platform+compiler
%  combinations (Sec.~}{)}{sec:install.requirements}, then once the
%compiler is selected in the platform file the build scripts will
%automatically adjust as necessary to produce a compatible
%\OOMMF\ build.  Otherwise, some adjustment of the platform
Some adjustment of the platform
configuration file, as described in the next section, may be
necessary.  Also, in many cases the compilers used to build 32-bit and
64-bit executables are different---if you encounter build problems,
double-check that the proper compiler is being used.

All of the \OOMMF\ downloads containing pre-built binaries are built for
use with thread-enabled \Tcl/\Tk.  You will need to build from source if
you want a non-threaded \OOMMF.  The build scripts will detect if the
\fn{tclsh} running the build procedure is non-threaded and will build
\OOMMF\ appropriately.

In all cases, use the platform configuration check described in the next
section to verify the compatibility of your \Tcl/\Tk\ and \OOMMF\
installs.

\subsection{Check Your Platform Configuration}
\index{platform!configuration}\label{sec:install.platformconfig}

After downloading and unpacking the \OOMMF\ software distribution, all
the \OOMMF\ software is contained in a subdirectory named \fn{oommf}.
Start a command line interface (a shell on \Unix, or a console
on \Windows), and change the
working directory\index{working~directory} to the
directory \fn{oommf}.
Find the \Tcl\ shell program installed as part of your \Tcl/\Tk\
installation.  In this manual we call the \Tcl\ shell program
\fn{tclsh}, but the actual name of the executable depends
on the release of \Tcl/\Tk\ and your platform type.  Consult
your \Tcl/\Tk\ documentation.

In the root directory of the \OOMMF\ distribution is a file
named \fn{oommf.tcl}.  It is the
\hyperrefhtml{bootstrap application}{bootstrap application (Ch.~}{)}{sec:cll}
which is used to launch all \OOMMF\ software.  With the command line
argument \cd{+platform},
it will print a summary of your
platform configuration when it is evaluated by \fn{tclsh}.
This summary describes your platform type, your \Cplusplus\ compiler,
and your \Tcl/\Tk\ installation.  As an example,
here is the typical output on a \MacOSX\ 10.9 system:
\begin{verbatim}
$ tclsh oommf.tcl +platform
<5426> oommf.tcl 1.2.0.6  info:
OOMMF release 1.2.0.6, snapshot 2015.03.25
Platform Name:    darwin
Tcl name for OS:  Darwin 13.4.0
C++ compiler:     /usr/bin/g++
 Version string:  Apple LLVM version 6.0 (clang-600.0.57) (based on
     LLVM 3.5svn) / Target: x86_64-apple-darwin13.4.0 / Thread model: posix
Shell details ---
 tclsh (running): /usr/bin/tclsh
     (links to /usr/bin/tclsh8.5)
     (links to /System/Library/Frameworks/Tcl.framework/Versions/8.5/tclsh8.5)
                      --> Version 8.5.9, 64 bit, threaded
 tclsh (OOMMF):   /usr/bin/tclsh8.5
                   --> Version 8.5.9, 64 bit, threaded
 filtersh:        /Users/xerxes/oommf/app/omfsh/darwin/filtersh
                   --> Version 8.5.9, 64 bit, threaded
 tclConfig.sh:    /System/Library/Frameworks/Tcl.framework/Versions/8.5/tclConfig.sh
                   --> Version 8.5.9
 wish (OOMMF):    /usr/bin/wish8.5
                   --> Version 8.5.9, Tk 8.5.9, 64 bit, threaded
 tkConfig.sh:     /System/Library/Frameworks/Tk.framework/Versions/8.5/tkConfig.sh
                   --> Tk Version 8.5.9
OOMMF threads:    Yes: Default thread count = 2
OOMMF API index:  20150129
Temp file directory: /var/folders/dy/srfj33512f51kc5knp_lph_r0000gp/T/
\end{verbatim}

If \cd{oommf.tcl +platform} doesn't print a summary similar to the
above, it should instead print an error message describing why it can't.
%For example, if your \Tcl\ installation is older than release 7.5, the
%error message will report that fact.
Follow any instructions provided
and repeat until \cd{oommf.tcl +platform} successfully prints a summary
of the platform configuration information.

\index{platform!names|(}
The first line of the example summary reports that \OOMMF\ recognizes
the platform by the name \cd{darwin}.  \OOMMF\ software recognizes many
of the more popular computing platforms, and assigns each a platform
name.  The platform name is used by \OOMMF\ in index and configuration
files and to name directories so that a single \OOMMF\ installation can
support multiple platform types.  If \cd{oommf.tcl +platform} reports
the platform name to be ``unknown'', then you will need to add some
configuration files to help \OOMMF\ assign a name to your platform type,
and associate with that name some of the key features of your computer.
See the section on
\hyperrefhtml{Managing \OOMMF\ platform names}
{``Managing \OOMMF\ platform names'' (Sec.~}{)}{sec:install.platformnames}
for further instructions.
\index{platform!names|)}

The second line reports the operating system version, which is mainly
useful to \OOMMF\ developers when fielding bug reports.  The third line
reports what \Cplusplus\ compiler will be used to build \OOMMF\ from its
\Cplusplus\ source code.  If you downloaded an \OOMMF\ release with
pre-compiled binaries for your platform, you may ignore this line.
Otherwise, if this line reports ``none selected'', or if it reports a
compiler other than the one you wish to use, then you will need to tell
\OOMMF\ what compiler to use.  To do that, you must edit the appropriate
configuration file for your platform.  Continuing the example above, one
would edit the file
\fn{config/platforms/darwin.tcl}\index{configuration~values}.  Editing
instructions are contained within the file.  On other platforms the name
\fn{darwin} in \fn{config/platforms/darwin.tcl} should be replaced with
the platform name \OOMMF\ reports for your platform.  For example, on a
32-bit Windows machine using an x86 processor, the corresponding
configuration file is \fn{config/platforms/wintel.tcl}.

\index{installation!Tcl/Tk|(}
The next group of lines describe the \Tcl\ configuration \OOMMF\ finds
on your platform.  The first couple of lines, ``tclsh (running)'',
describe the \Tcl\ shell running the oommf.tcl script.  After that, the
``tclsh (\OOMMF)'' subgroup describes the \Tcl\ shell that \OOMMF\ will
launch when it needs to run \Tcl\ scripts.  If the \OOMMF\ binaries
have been built, then there will also be a \cd{filtersh} subgroup, which
describes the augmented \Tcl\ shell used to run many of the
\OOMMF\ support scripts.  All of these shells should report the same
version, bitness, and threading information.  If \OOMMF\ can't find
\cd{tclsh}, or if it finds the wrong one, you can correct this by
setting the environment variable
OOMMF\_TCLSH\index{environment~variables!OOMMF\_TCLSH} to the absolute
location of \cd{tclsh}.  (For information about setting environment
variables, see your operating system documentation.)

Following the \Tcl\ shell information, the \cd{tclConfig.sh} lines
report the name of the configuration file installed as part of \Tcl, if
any.  Conventional \Tcl\ installations on Unix systems and within the
Cygwin environment\index{platform!Windows!Cygwin~environment} on
\Windows\ have such a file, usually named \fn{tclConfig.sh}.  The
\Tcl\ configuration file records details about how \Tcl\ was built and
where it was installed.  On \Windows\ platforms, this information is
recorded in the
\Windows\ registry\index{platform!Windows!no~Tcl~configuration~file}, so
it is normal to have \cd{oommf.tcl +platform} report ``none found''.  If
\cd{oommf.tcl +platform} reports ``none found'', but you know that an
appropriate \Tcl\ configuration file is present on your system, you can
tell \OOMMF\ where to find the file by setting the environment variable
OOMMF\_TCL\_CONFIG\index{environment~variables!OOMMF\_TCL\_CONFIG} to
its absolute filename.  In unusual circumstances, \OOMMF\ may find a
\Tcl\ configuration file which doesn't correctly describe your
\Tcl\ installation.  In that case, use the environment variable
OOMMF\_TCL\_CONFIG to instruct \OOMMF\ to use a different file that you
specify, and, if necessary, edit that file to include a correct
description of your \Tcl\ installation.

Next, the \cd{oommf.tcl +platform} reports similar information about the
\cd{wish} and \Tk\ configuration.  The environment variables
\cd{OOMMF\_TK\_CONFIG}\index{environment~variables!OOMMF\_TK\_CONFIG}
and \cd{OOMMF\_WISH}\index{environment~variables!OOMMF\_WISH} may be
used to tell \OOMMF\ where to find the \Tk\ configuration file and the
\fn{wish} program, respectively.

\index{parallelization|(}
Following the \Tk\ information are some lines reporting ``thread'' build
and run status. Threads are used by \OOMMF\ to implement parallelism in
the Oxs (\cd{oxsii} and \cd{boxsi}) 3D solvers on
multi-processor/multi-core shared memory machines.  In order to build or
run a parallel version of OOMMF, you must have a thread-enabled version
of \Tcl.  The \Tcl\ thread status is indicated on the first thread
status line.  If \Tcl\ is thread enabled, then the default \OOMMF\ build
process will create a threaded version of \OOMMF.  You can override this
behavior if you wish to build a non-parallel version of \OOMMF\ by
editing the
\cd{oommf\_threads}\index{configuration~values!oommf\_threads} value in
the \fn{config/platforms/} file for your platform.

If \Tcl\ and \OOMMF\ threads are enabled, then the default number of
threads run by the Oxs solvers is also reported.  (This value may vary
between machines, depending on the number of processors in the
machine.)  You can change this by setting (in order of increasing
precedence) the
\cd{oommf\_thread\_count}\index{configuration~values!oommf\_thread\_count}
value in the installation-wide \fn{config/options} file, the
\cd{thread\_count}\index{configuration~values!thread\_count} value in
the \fn{config/platforms/} file for your platform, via the environment
variable
\cd{OOMMF\_THREADS}\index{environment~variables!OOMMF\_THREADS}, or by
the \cd{oxsii}/\cd{boxsi} command line option \cd{-threads}.

By default, \OOMMF\ sets no upper limit on the number of threads you
may run in \cd{oxsii} or \cd{boxsi}.  However, performance is degraded
if you run more threads than available cpu cores.  To protect against
this, or to limit resource use on a shared machine, you may wish to
set a hard limit on the maximum number of threads per \cd{oxsii} or
\cd{boxsi} instance.  This can be done by setting (in order of
increasing precedence) the environment variable
\cd{OOMMF\_THREADLIMIT}\index{environment~variables!OOMMF\_THREADLIMIT},
the \cd{thread\_limit}\index{configuration~values!thread\_limit} value
in the \fn{config/platforms/} file for your platform, or the
\cd{oommf\_thread\_limit}\index{configuration~values!oommf\_thread\_limit}
value in the \fn{config/options} file.  (Note the precedence order is
reversed compared to that for the default thread count.)  If a limit
is set then that value is displayed in the threads line of the
\cd{oommf.tcl +platform} output.

If NUMA\index{NUMA} support is provided on your platform (see
\hyperrefhtml{below}{``Parallelization,'' Sec.~}{
  below}{sec:install.parallel}), then the following \cd{oommf.tcl
  +platform} output line will indicate whether or not the build process
will create NUMA-aware Oxs solvers.\index{parallelization|)}

After the thread and NUMA information, \cd{oommf.tcl +platform}
reports the directory that \OOMMF\ will use to write temporary
files\index{temporary~files}.  This directory is used, for example, to
transfer magnetization data from the micromagnetic solvers to the
\app{mmDisp} display module.  You must have write access to this
directory.  It needs to have enough space to manage the dataflows of
your simulations.  It is also beneficial if this directory is local to
the processors performing the calculations.  If you don't like the
\OOMMF\ default, you may change it via the
\cd{path\_directory\_temporary}\index{configuration~values!path\_directory\_temporary}
setting in the \fn{config/platforms/} file for your platform.  Or you
can set the environment variable
\cd{OOMMF\_TEMP}\index{environment~variables!OOMMF\_TEMP}, which will
override all other settings.

If any environment variables relevant to \OOMMF\ are set, then
\cd{oommf.tcl +platform} will report these next, followed finally by any
warnings about possible problems with your \Tcl/\Tk\ installation, such
as if you are missing important header files.

If \cd{oommf.tcl +platform} indicates problems with your \Tcl/\Tk\
installation,
it may be easiest to re-install \Tcl/\Tk\, taking care to perform a
conventional installation.  \OOMMF\ deals best with conventional
\Tcl/\Tk\ installations.  If you do not have the power to re-install
an existing broken \Tcl/\Tk\ installation (perhaps you are not
the sysadmin of your machine), you might still install your own
copy of \Tcl/\Tk\ in your own user space.  In that case, if your
private \Tcl/\Tk\ installation makes use of shared libraries,
take care that you do whatever is necessary on your platform to
be sure that your private \fn{tclsh} and \fn{wish}
find and use your private
shared libraries instead of those from the system \Tcl/\Tk\ installation.
This might involve setting an environment variable (such as
LD\_LIBRARY\_PATH\index{environment~variables!LD\_LIBRARY\_PATH} on
\Unix\, or PATH\index{environment~variables!LD\_LIBRARY\_PATH} on
\Windows).
If you use a private \Tcl/\Tk\ installation, you also want to be sure
that there are no environment variables like
TCL\_LIBRARY\index{environment~variables!TCL\_LIBRARY}
or TK\_LIBRARY\index{environment~variables!TK\_LIBRARY}
that still refer to the system \Tcl/\Tk\ installation.
\index{installation!Tcl/Tk|)}

\subsubsection{Additional Configuration Issues on \Windows}
\index{platform!Windows!configuration|(}
A few other configurations should be checked on \Windows\ platforms.
\index{platform!Windows!file~path~separator|(}
First, note that absolute filenames on \Windows\ makes use of the
backslash (\bs) to separate directory names.  On \Unix\ and
within \Tcl\ the forward slash (\fs) is used to separate directory
names in an absolute filename.  In this manual we usually use the
\Tcl\ convention of forward slash as separator.  In portions of the
manual pertaining only to MS \Windows\ we use the backslash as
separator.  There may be instructions in this manual which do not
work exactly as written on \Windows\ platforms.  You may need to
replace forward slashes with backward slashes in pathnames when
working on \Windows.
\index{platform!Windows!file~path~separator|)}

\index{platform!Windows!hosts~file|(}
\OOMMF\ software needs networking support that recognizes
the host name \cd{localhost}.  It may be necessary
to edit a file which records that \cd{localhost} is a synonym
for the loopback interface (127.0.0.1).  If a file named \fn{hosts}
exists in your system area (for example, \fn{C:\bs Windows\bs hosts}),
be sure it includes an entry mapping 127.0.0.1 to \cd{localhost}.
If no \fn{hosts} file exists, but a \fn{hosts.sam} file exists,
make a copy of \fn{hosts.sam} with the name \fn{hosts}, and edit
the copy to have the \fn{localhost} entry.
\index{platform!Windows!hosts~file|)}

The directory that holds the \fn{tclsh} and \fn{wish} programs also
holds several {\fn{*.dll}} files that \OOMMF\ software needs to find to
run properly.  Normally when the \hyperrefhtml{\OOMMF\ bootstrap
  application}{\OOMMF\ bootstrap application (Ch.~}{)}{sec:cll} or
\hyperrefhtml{{\bf mmLaunch}}{{\bf mmLaunch} (Ch.~}{)}{sec:mmlaunch} is
used to launch \OOMMF\ programs, they take care of making sure the
necessary {\fn{*.dll}} files can be found.  As an additional measure,
you might want to add the directory which holds the \fn{tclsh} and
\fn{wish} programs to the list of directories stored in the
PATH\index{environment~variables!PATH} environment variable.  All the
directories in the PATH are searched for {\fn{*.dll}} files needed when
starting an executable.  \index{platform!Windows!configuration|)}

\subsection{Compiling and Linking}\label{sec:install.compile}

If you downloaded a distribution with pre-compiled executables, you may
skip this section.

When building \OOMMF\ software from source code, be sure the
\Cplusplus\ compiler reported by \cd{oommf.tcl +platform} is properly
configured.  In particular, if you are running on a \Windows\ system,
please read carefully the notes in \html{the} \hyperrefhtml{Advanced
  Installation}{Advanced Installation, Sec.~}{,}{sec:install.windows}
\html{section} pertaining to your compiler.

The compiling and linking of the \Cplusplus\ portions of
\OOMMF\ software are guided by the application
\hyperrefhtml{pimake}{pimake (Sec.~}{)}{sec:pimake}
\index{application!pimake}
(``Platform Independent Make'') which
is distributed as part of the \OOMMF\ release.
To begin building \OOMMF\ software with \fn{pimake}, first change
your working directory\index{working~directory}
to the root directory of the \OOMMF\ distribution:
\begin{verbatim}
cd .../path/to/oommf
\end{verbatim}

If you unpacked the new \OOMMF\ release into a directory \fn{oommf}
which contained an earlier \OOMMF\ release,
use \fn{pimake} to build the target \cd{upgrade}
to clear away any source code files which were
in a former distribution but are not part of the latest distribution:
\begin{verbatim}
tclsh oommf.tcl pimake upgrade
\end{verbatim}

Next, build the target \cd{distclean} to clear away any old executables
and object files which are left behind from the compilation of the
previous distribution:
\begin{verbatim}
tclsh oommf.tcl pimake distclean
\end{verbatim}

Next, to build all the \OOMMF\ software, run \fn{pimake} without
specifying a target:
\begin{verbatim}
tclsh oommf.tcl pimake
\end{verbatim}
On some platforms, you cannot successfully compile \OOMMF\ software if
there are \OOMMF\ programs running.  Check that all \OOMMF\ programs
have terminated (including those in the background) before trying to
compile and link \OOMMF.

When \fn{pimake} calls on a compiler or other software development
utility, the command line is printed, so that you may monitor the build
process.
Assuming a proper configuration for your platform, \fn{pimake} should be
able to compile and link all the \OOMMF\ software without error.  If
\fn{pimake} reports errors, please first consult
\hyperrefhtml{Troubleshooting}{Troubleshooting (Ch.~}{)}{sec:trouble}
to see if a fix is already documented.
If not, please send both the {\em complete} output
from \fn{pimake} and the output from \cd{oommf.tcl +platform}
to the \OOMMF\ developers when you e-mail to ask for help.

\subsection{Installing}\label{sec:install.installing}
% Note: LaTeXML uses \label value to set file names.
The current \OOMMF\ release does not support an installation procedure.
For now, simply run the executables from the directories in which they
were unpacked/built.

\subsection{Using \OOMMF\ Software}\label{sec:install.using}
% Note: LaTeXML uses \label value to set file names.
To start using \OOMMF\ software, run the
\hyperrefhtml{\OOMMF\ bootstrap application}{\OOMMF\ bootstrap
application (Ch.~}{)}{sec:cll}. This may be launched from the command
line interface:
\begin{verbatim}
tclsh oommf.tcl
\end{verbatim}

If you prefer, you may launch the \OOMMF\ bootstrap application
\fn{oommf.tcl} using whatever graphical ``point and click''
interface your operating system provides.  By default, the \OOMMF\
bootstrap application will start up a copy of the \OOMMF\ application
\hyperrefhtml{\app{mmLaunch}}{\app{mmLaunch} (Ch.~}{)}{sec:mmlaunch} in a
new window.

If you publish material created with the aid of \OOMMF, please refer to
\hyperrefhtml{Credits}{Credits (Ch.~}{)}{sec:credits}
for citation information.


\subsection{Reporting Problems}\label{sec:install.reporting}
% Note: LaTeXML uses \label value to set file names.
If you encounter problems when installing or using \OOMMF, please report
them to the \OOMMF\ developers.  The \cd{oommf.tcl +platform} command
has been designed in large part to help \OOMMF\ developers debug
installation problems, so \textbf{PLEASE} be sure to include the
complete output from \cd{oommf.tcl~+platform} in your report.  See also
the section on
\hyperrefhtml{troubleshooting}{troubleshooting (Ch.~}{)}{sec:trouble}
for additional instructions.

\section{Advanced Installation}\label{sec:install.advanced}

The following sections provide instructions for some additional
installation options.

\subsection{Reducing Disk Space Usage}\label{sec:install.reducedisk}

\index{requirement!disk~space|(}
To delete the intermediate files created when building the \OOMMF\
software from source code, use
\hyperrefhtml{pimake}{pimake (Sec.~}{)}{sec:pimake}
to build the target
\cd{objclean} in the root directory of the \OOMMF\ distribution.
\begin{verbatim}
tclsh oommf.tcl pimake objclean
\end{verbatim}
Running your platform \fn{strip} utility on the \OOMMF\ executable files
should also reduce their size somewhat.
\index{requirement!disk~space|)}

\subsection{Local Customizations}\label{sec:install.custom}
\index{customize}
\OOMMF\ software supports local customization of some of its
features. The files controlling these options are
\fn{config/options.tcl}\index{options.tcl} and
\fn{config/platforms/<platform>.tcl},\index{platform.tcl} where
``\fn{<platform>}'' is the identifier for the platform, e.g.,
\fn{windows-x86\_64} for \Windows, \fn{linux-x86\_64} for \Linux, and
\fn{darwin} for \MacOSX. Although you may edit these files directly, it
is better to create a subdirectory named \fn{local} and
create a file with the same name containing your changes. (Separating
your changes from the distributed versions of these files makes it
easier to port your changes to new versions of \OOMMF.)  For example, to
change the settings in the \fn{options.tcl} file, create the file
\fn{config/local/options.tcl}. To modify the platform file on \Windows,
create the file \fn{config/platforms/local/windows-x86\_64.tcl}.

The \cd{options.tcl} and platform files are \Tcl\ script files, and you
may use any of the standard \Tcl\ commands and constructs inside
them. There are also \OOMMF-specific commands available in these files,
in particular \cd{Oc\_Option} for \cd{options.tcl}, and \cd{\$config
  SetValue} for platform files.

The \fn{options.tcl} file controls settings which are platform agnostic.
For example, this file sets the default network port for the
\hyperrefhtml{host service directory application}{host service
  directory application (Ch.~}{)}{sec:arch}, the location of
\hyperrefhtml{\app{Oxs}}{\app{Oxs} log files (Ch.~}{),}{sec:oxs}
\HTMLoutput{log files,} and the width of curves in
\hyperrefhtml{\app{mmGraph}}{\app{mmGraph} (Ch.~}{)}{sec:mmgraph}.
Editing instructions are provided in \fn{options.tcl}. As an example,
the setting
\begin{verbatim}
Oc_Option Add mmGraph Ow_GraphWin default_curve_width 1
\end{verbatim}
in \fn{options.tcl} can be overridden in \fn{local/options.tcl} with
\begin{verbatim}
Oc_Option Add mmGraph Ow_GraphWin default_curve_width 2
\end{verbatim}
if wider lines are desired. The \fn{options.tcl} file also controls some
build and optimization controls, as discussed
\HTMLoutput{in the }
\hyperrefhtml{next section}{below (Sec.~}{)}{sec:install.optimize}.

The platform configuration files control settings that are platform
dependent, such as the location for temporary files, compiler selection,
and compiler and linker search paths. These files contain a block of
code demarcated by the string \verb+LOCAL CONFIGURATION+ that lists
end-user configuration options. These controls are all commented out
(i.e., deactivated by a leading \verb+#+ character, which is the comment
marker in \Tcl), so you can copy this entire block into your
\fn{local/<platform>.tcl} file and uncomment only the options you want
to change. For example, if you are using
\htmladdnormallinkfoot{MacPorts}{https://www.macports.org}
on \MacOSX, then you
may want to edit your \fn{config/platforms/local/darwin.tcl} file
to have the lines
\begin{verbatim}
## EXTERNAL PACKAGE SUPPORT:
## Extra include directories for compiling:
$config SetValue program_compiler_extra_include_dirs /opt/local/include
#
## Extra directories to search for libraries.
$config SetValue program_linker_extra_lib_dirs /opt/local/lib
\end{verbatim}
to allow the compiler and linker to search in the MacPorts default
install area.

There are a number of environment
variables\index{environment~variables!customization} that can be set to
override the values in the above configuration files, including
\begin{description}
\item[OOMMF\_CPP:] Two item list specifying the \Cplusplus\ compiler and
  the option to compile only (no link); e.g., for a Bourne compatible shell
\begin{verbatim}
  export OOMMF_CPP=g++ -c
\end{verbatim}
selects the GNU \Cplusplus\ compiler. To select the Visual C++ compiler
at the \Windows\ command prompt, use
\begin{verbatim}
  set OOMMF_CPP=cl /c
\end{verbatim}
Overrides the \fn{<platform>.tcl} \cd{program\_compiler\_c++\_override}
setting.
\item[OOMMF\_HOSTPORT:] Host server listen port. Overrides the
  \fn{options.tcl} ``\cd{Net\_Host port}'' value.
\item[OOMMF\_OUTDIR:] Output file directory for \app{Oxs} runs;
  overrides default which is the directory holding the run \fn{.mif}
  simulation specification file.
\item[OOMMF\_OXSLOGDIR:] Directory for \app{Oxs} log files. Overrides
  \fn{options.tcl} setting ``\cd{OxsLogs directory}''.
\item[OOMMF\_TEMP:] Directory for temporary files. Overrides auto-selection.
\item[OOMMF\_THREADS:] Default number of execution threads for \app{Oxs}
  runs. Overrides \fn{<platform>.tcl} \cd{thread\_count} setting.
\end{description}
Be aware that environment variable settings hold only for applications
launched from the given shell instance, unless the value is set in the
shell start-up configuration.

Where applicable, command line options override both configuration file
and environment variable settings. For example, \app{oxsii} and
\app{boxsi} support the \cd{-threads} option that specifies the number
of execution threads.

The \OOMMF\ installation can also be customized by the installation of
third party and experimental extensions to the \hyperrefhtml{\OOMMF\
  eXtensible Solver (Oxs)}{\OOMMF\ eXtensible Solver (Oxs)
  (Ch.~}{)}{sec:oxs}. These are maintained using the
\hyperrefhtml{\app{oxspkg}}{\app{oxspkg} (Sec.~}{)}{sec:oxspkg} package
  management tool.


\subsection{Optimization}\label{sec:install.optimize}\index{optimization}
In the interest of successful compilation of a usable software package
``out of the box,'' the default configuration for \OOMMF\ does not
stretch to maximize optimization.  However, the configuration files
\fn{config/options.tcl} and \fn{config/platforms/<platform>.tcl} include
optimization controls.  In the former,
\begin{verbatim}
Oc_Option Add * Platform optlevel 2
\end{verbatim}
selects a relatively high but robust level of optimization, equivalent
typically to a compiler option of ``-O2'' or similar. You may
try increasing \cd{optlevel} to 3. If you are familiar with your
compiler's command line options, then for finer control you can
experiment with the \fn{<platform>.tcl} file options
\begin{duplex}
\item \verb!program_compiler_c++_remove_flags!
\item[and]\html{\\}
\item \verb!program_compiler_c++_add_flags!
\end{duplex}
The remove flags control selects options to be removed from the compiler
command line. The setting value should be a \Tcl\ list of regular
expressions.  (Refer to the
\htmladdnormallinkfoot{\Tcl\ documentation}{https://www.tcl-lang.org/man/}
on the regexp command for details.) The add flags value is a list of
values to append to the compile command (applied after the remove flags
request, if any). An extended discussion of these controls can be found
in the ``Configuration Files'' section of the
\htmladdnormallinkfoot{\textit{\OOMMF\ Programming
    Manual}}{https://math.nist.gov/oommf/doc/}.

\subsection{Parallelization}%
\label{sec:install.parallel}\index{parallelization|(}
The \OOMMF\ Oxs 3D solvers (\cd{oxsii} and \cd{boxsi}) can be built
thread-enabled to allow parallel processing on
multi-processor/multi-core machines.  In order to build and run a
parallel version of \OOMMF, you must have a thread-enabled version of
\Tcl.  Most standard binary releases of \Tcl\ today are thread-enabled,
so \OOMMF\ releases that include pre-built executables are built
thread-enabled.  If you build \OOMMF\ from source, then by default
\OOMMF\ will be built thread-enabled if your \Tcl\ is thread-enabled.
As explained earlier, you can check thread build status with the
\cd{tclsh oommf.tcl +platform} command.  If you want to force a
non-threaded build of \OOMMF, then edit the \fn{config/platforms/} file
for your platform.  In the section labeled \cd{LOCAL CONFIGURATION}, you
will find a line that looks like
\begin{verbatim}
# $config SetValue oommf_threads 0
\end{verbatim}
Uncomment this line (i.e., remove the leading `\verb+#+' character) to
force a non-threaded build.  Then run
\begin{verbatim}
tclsh oommf.tcl pimake distclean
tclsh oommf.tcl pimake
\end{verbatim}
from the \OOMMF\ root directory to create a fresh build.

You can use the \cd{tclsh oommf.tcl +platform} command to see the
default number of compute threads that will be run by the Oxs 3D solver
programs \cd{oxsii} and \cd{boxsi}.  You can modify the default as
explained in the \hyperrefhtml{Check Your Platform
  Configuration}{Platform Configuration
  (Sec.~}{)}{sec:install.platformconfig} and \hyperrefhtml{Local
  Customizations}{Local Customizations (Sec.~}{)}{sec:install.custom}
sections, or you can override the default at run time via the command
line option \cd{-threads} to \cd{oxsii} and \cd{boxsi}.

\index{NUMA|(}Some multi-processor machines have a non-uniform memory
architecture (NUMA), which means that although each processor can access
all of system memory, some parts of memory can be accessed faster than
others.  Typically this is accomplished by dividing the system memory
and processors into ``nodes.''  Memory accesses within a node are faster
than accesses between nodes, and depending on the architecture access
latency and bandwidth may be different between different node pairs.
Examples of machines with NUMA include some multi-processor AMD Opteron
and Intel Xeon boxes.

Computer programs such as \OOMMF\ can run on NUMA machines without
making any special allowances for the memory architecture.  However, a
program that is written to take advantage of the faster local
(intra-node) memory accesses can sometimes run significantly faster.
\OOMMF\ contains NUMA-aware code, but this code is highly operating
system specific.  At present, \OOMMF\ can be built with NUMA support
only on Linux (32- and 64-bit) systems.  To do this, you must install
the operating system NUMA support packages ``numactl'' and
``numactl-devel''.  The names may vary somewhat between Linux
distributions, but the first typically includes the executable
\fn{numactl} and the second includes the header file \fn{numa.h}.  Once
the numactl package is installed, you can run the command
\begin{verbatim}
numactl --hardware
\end{verbatim}
to get an overview of the memory architecture on your machine.  If this
shows you have only one node, then there is no advantage to making a
NUMA-aware build of \OOMMF.

The next step is to edit the \fn{config/platforms} for your platform.
For example, on a 64-bit Linux box this file is
\fn{config/platforms/linux-x86\_64.tcl}.
In the section labeled \cd{LOCAL CONFIGURATION}, find the line
\begin{verbatim}
# $config SetValue use_numa 1
\end{verbatim}
Edit this to remove the leading `\verb+#+' character.  Alternatively
(and, actually, preferably), create a \fn{local} subdirectory and make a
local configuration file with the same platform name; e.g.,
\fn{config/platforms/local/linux-x86\_64.tcl} on a 64-bit Linux machine.  Add
the line
\begin{verbatim}
$config SetValue use_numa 1
\end{verbatim}
to this file.  (The advantage of using a \fn{config/platforms/local}
file is that you can make changes without modifying the original \OOMMF\
source code, which makes it easier to port your local changes to future
releases of \OOMMF.)  If this is done correctly, then the command
`\cd{tclsh oommf.tcl +platform}' will show that NUMA support is enabled.
Then simply run `\cd{tclsh oommf.tcl pimake distclean}' and
`\cd{tclsh oommf.tcl pimake}' from the \OOMMF\ root directory to build a
NUMA-aware version of \OOMMF.

To activate the NUMA-aware code, you must specify the \cd{-numanodes}
option on the \cd{oxsii}/\cd{boxsi} command line, or set the the
environment variable
\cd{OOMMF\_NUMANODES}\index{environment~variables!OOMMF\_NUMANODES}.
Check the \hyperrefhtml{Oxs documention}{Oxs documentation
(Ch.~}{)}{sec:oxs} for details.
\index{NUMA|)}\index{parallelization|)}

\subsection{Managing \OOMMF\ Platform Names}\label{sec:install.platformnames}
\index{platform!names|(}

\OOMMF\ software classifies computing platforms into different types
using the scripts in the directory \fn{config/names} relative to the
root directory of the \OOMMF\ distribution.
Each type of computing platform is assigned a unique name.
These names are used as directory names
and in index and configuration files so that a single
\OOMMF\ installation may contain platform-dependent sections for many
different types of computing platforms.

To learn what name \OOMMF\ software uses to refer to your computing
platform, run
\begin{verbatim}
tclsh oommf.tcl +platform
\end{verbatim}
in the \OOMMF\ root directory.

\subsubsection{Changing the name \OOMMF\ assigns to your platform}
First, use
\hyperrefhtml{pimake}{pimake (Sec.~}{)}{sec:pimake}
to build the target \cd{distclean} to
clear away any compiled executables built using the old platform
name.
\begin{verbatim}
tclsh oommf.tcl pimake distclean
\end{verbatim}
Then, to change the name \OOMMF\ software uses to describe your platform from
\cd{foo} to \cd{bar}, simply rename the file
\begin{quote}
\fn{config/names/foo.tcl}
\hspace{1em} to \hspace{1em}
\fn{config/names/bar.tcl}
\end{quote}
and
\begin{quote}
\fn{config/platforms/foo.tcl}
\hspace{1em} to \hspace{1em}
\fn{config/platforms/bar.tcl}.
\end{quote}
After renaming your platform type, you should recompile your executables
using the new platform name.

\subsubsection{Adding a new platform type}
If \cd{oommf.tcl +platform} reports the platform name
\cd{unknown}, then none of the scripts in \fn{config/names/}
recognizes your platform type.  As an example, to add the platform
name \cd{foo} to \OOMMF's vocabulary of platform names, create the
file \fn{config/names/foo.tcl}.  The simplest way to proceed is to
copy an existing file in the directory \fn{config/names} and edit it
to recognize your platform.

The files in \fn{config/names} include \Tcl\ code like this:

\begin{verbatim}
  Oc_Config New _ \
    [string tolower [file rootname [file tail [info script]]]] {
      # In this block place the body of a Tcl proc which returns 1
      # if the machine on which the proc is executed is of the
      # platform type identified by this file, and which returns 0
      # otherwise.
      #
      # The usual Tcl language mechanism for discovering details
      # about the machine on which the proc is running is to
      # consult the global Tcl variable 'tcl_platform'.  See the
      # existing files for examples, or contact the OOMMF
      # developers for further assistance.
  }
\end{verbatim}

After creating the new platform name file \fn{config/names/foo.tcl}, you
also need to create a new platform file \fn{config/platforms/foo.tcl}.
A reasonable starting point is to copy the file
\fn{config/platforms/unknown.tcl} for editing.  Contact the \OOMMF\
developers for assistance.

Please consider contributing your new platform recognition and
configuration files to the \OOMMF\ developers for inclusion in
future releases of \OOMMF\ software.

\subsubsection{Resolving platform name conflicts}
If the script \cd{oommf.tcl +platform} reports ``Multiple platform names are
compatible with your computer'', then there are multiple files in the
directory \fn{config/names/} that return 1 when run on your computer.
For each compatible platform name reported, edit the corresponding
file in \fn{config/names/} so that only one of them returns 1.
Experimenting using \fn{tclsh} to probe the \Tcl\ variable
\cd{tcl\_platform} should assist you in this task.  If that fails, you
can explicitly assign a platform type corresponding to your computing
platform by matching its hostname.  For example, if your machine's
host name is {\tt foo.bar.net}:
\begin{verbatim}
  Oc_Config New _ \
    [string tolower [file rootname [file tail [info script]]]] {
      if {[string match foo.bar.net [info hostname]]} {
          return 1
      }
      # Continue with other tests...
  }
\end{verbatim}

Contact the \OOMMF\ developers if you need further assistance.
\index{platform!names|)}

\section{Platform Specific Installation Issues}\label{sec:install.platformissues}
The installation procedure discussed in the previous sections applies to
all platforms (\Unix, \Windows, \MacOSX).  There are, however, some
details which pertain only to a particular platform.  These issues are
discussed below.

\index{platform!Unix!configuration|(}
\subsection{\Unix\ Configuration}\label{sec:install.unixconfig}
% Note: LaTeXML uses \label value to set file names.
\subsubsection{Missing \Tcl/\Tk\ files}
The basic installation procedure should be sufficient to install \OOMMF\
on most \Unix\ systems.  Sometimes, however, the build will fail due to
missing \Tcl\ header files (\fn{tcl.h}, \fn{tk.h}) or libraries (e.g.,
\fn{libtcl.so}, \fn{libtk.so}).  This problem can usually be solved by
installing a ``development'' version of \Tcl/\Tk, which may be found on
the operating system installation disks, or may be available from the
system vendor.  There are also binary releases of \Tcl/\Tk\ for a number
of systems available from ActiveState, under the name
\htmladdnormallinkfoot{ActiveTcl}{https://www.activestate.com/products/tcl/}.
Alternatively, one may download the sources for \Tcl\ and \Tk\ from the
\htmladdnormallinkfoot{Tcl Developer
Xchange}{http://purl.org/tcl/home/}, and build and install \Tcl/\Tk\
from source.  The \Tcl/\Tk\ build follows the usual \Unix\
\fn{configure}, \fn{make}, \fn{make install} build convention.

\subsubsection{Compiler Optimization Options}
On most systems, \OOMMF\ builds by default with relatively unaggressive
compiler optimization options.  As discussed earlier
(\html{under }\hyperrefhtml{Optimization)}{``Optimization,''
Sec.~}{)}{sec:install.optimize}, you may edit the appropriate
\fn{oommf/config/platforms/} file to change the default compilation options.
However, on some common systems (e.g., Linux, some BSD variants) \OOMMF\
will try to deduce the hardware architecture (i.e., the CPU subtype,
such as Pentium 3 vs. Pentium 4) and apply architecture-specific options
to the compile commands.  This is probably what you want if \OOMMF\ is
to be run only on the system on which it was built, or if it is run on a
homogeneous cluster.  If, instead, you intend to run \OOMMF\ on a
heterogeneous cluster you may need to restrict the compiler options to
those supported across your target machines.  In that case, open the
appropriate configuration file in the \fn{oommf/config/platforms/}
directory, and look for the lines
\begin{verbatim}
    # You can override the GuessCPU results by directly setting or
    # unsetting the cpuopts variable, e.g.,
    #
    #    set cpuopts [list -march=skylake]
    # or
    #    unset cpuopts
    #
\end{verbatim}
Uncomment either the ``unset cpuopts'' line to make a generic build, or
else edit the ``set cpuopts'' line to an appropriate common-denominator
architecture and uncomment that line.

In a similar vein, some compilers support a ``-fast'' switch, which
usually creates an architecture-specific executable.  The same
considerations apply in this case.

An advanced alternative would be to define separate \OOMMF\
``platforms'' for each CPU subtype in your cluster.  At a minimum, this
would involve creating separate platform name files in
\fn{oommf/config/names/} for each subtype, and then making copies of the
appropriate \fn{oommf/config/platforms} file for each new platform.  The
platform name files would have to be written so as to reliably detect
the CPU subtype on each machine.  See \hyperrefhtml{``Managing \OOMMF\
platform names''} {``Managing \OOMMF\ platform names''
(Sec.~}{)}{sec:install.platformnames} for details on creating platform name
files.  \index{platform!Unix!configuration|)}

\subsubsection{Portland Group pgCC compiler on Linux}
The platform build scripts for Linux,
\fn{oommf/config/platforms/lintel.tcl} (32-bit) and
\fn{oommf/config/platforms/linux-x86\_64.tcl} (64-bit) contain sections
supporting the Portland Group pgCC compiler.  Non-threaded builds of
OOMMF using this compiler run fine, but threaded builds segfault when
running \hyperrefhtml{\fn{Oxsii}/\fn{Boxsi}} {\fn{Oxsii}/\fn{Boxsi}
 (Ch.~}{)}{sec:oxs}.  The source of this problem is not known at this
time.


\subsection{\MacOSX\ Configuration}\label{sec:install.macosx}
\index{platform!macOS!configuration|(}
The build procedure for \MacOSX\ is the same as for \Unix.  The
platform name is ``darwin''.  If the \hyperrefhtml{platform
configuration check}{platform configuration check
(Sec.~}{)}{sec:install.platformconfig} does not find a \Cplusplus\ compiler,
then you will have to install one.  One option is the Xcode command line
developer tools provided by Apple.  You can install these from a
Terminal window via the command
\begin{verbatim}
xcode-select --install
\end{verbatim}
You should run this command even if you install the full Xcode IDE.
Refer to your system documentation for details.
\index{platform!macOS!configuration|)}

\index{platform!Windows!configuration|(}
\subsection{Microsoft \Windows\ Options}\label{sec:install.windows}

This section lists installation options for Microsoft \Windows.

\subsubsection{Using Microsoft Visual \Cplusplus}
\index{platform!Windows!Microsoft~Visual~C++}
If you are building \OOMMF\ software from source using the Microsoft
Visual \Cplusplus\ command line compiler, \fn{cl.exe}, it is necessary to set up
the path and some environment variables before running the compiler.
There is a batch file distributed with Visual \Cplusplus\ that you can run to do
this.  The name of the file varies between Visual \Cplusplus\ releases, but for
example may be \fn{vcvarsall.bat} or \fn{setenv.cmd}.  For 64-bit builds
you may need to include the ``amd64'' option on the batch file command
line.  You may want to set up your system so this batch file gets run
automatically when you open a command window.  See your compiler and
system documentation for details.

\subsubsection{Using MinGW g++}
\index{platform!Windows!MinGW~g++}

Both 32-bit and 64-bit builds are supported using the MinGW ports of
g++.  (The 32-bit and 64-bit versions of g++ are separate downloads.)
Use a standard Windows \Tcl/\Tk, such as the
\htmladdnormallinkfoot{ActiveTcl}{https://www.activestate.com/products/tcl/}
release from ActiveState.  You will also need to edit the appropriate
platform file to select g++ as the compiler.  If you are using a 32-bit
\Tcl/\Tk\ and g++, then the platform file is
\fn{oommf\bs config\bs platforms\bs wintel.tcl}.  For 64-bit
\Tcl/\Tk\ and g++ the platform file is
\fn{oommf\bs config\bs platforms\bs windows-x86\_64.tcl}.

\subsubsection{Using the Cygwin toolkit}\label{par:install.cygwin}
\index{platform!Windows!Cygwin~environment}
The \htmladdnormallinkfoot{Cygwin Project}{http://www.cygwin.com/} is
a free port of the GNU development environment to \Windows, which
includes the GNU C++ compiler g++ and X11.  To build OOMMF within the
Cygwin environment, start up a Cygwin or Cygwin64 shell and follow the
usual \Unix\ build procedure.  The platform name will be \fn{cygtel}
or \fn{cygwin-x86\_64}, according to whether you are running a 32- or
64-bit Cygwin \fn{tclsh}, respectively.  The resulting \OOMMF\ build
requires the Cygwin environment to run, so it will need to be launched
from a Cygwin shell.  Moreover, \OOMMF\ on Cygwin uses X11 as the
windowing interface, so you will need to have the Cygwin port of X11
installed; typically \OOMMF\ will be started from an X11 xterm or
equivalent.  Of course, you will also need the \Tcl\ and \Tk\ packages
installed (called \fn{tcl} and \fn{tcl-tk}, respectively, by the
Cygwin package manager).  To build \OOMMF\ from source you will
need the \fn{gcc-g++}, \fn{tcl-devel}, and \fn{tcl-tk-devel}
packages and dependencies.

If you get errors saying a child process couldn't be forked (typically
with either ``resource temporarily unavailable'' or ``Loaded to
different address'' error messages), then follow this procedure:
\begin{enumerate}
\item Exit all Cygwin processes
\item Use Windows Explorer or a Windows command shell to launch
  \fn{c:\bs cygwin\bs bin\bs ash.exe}
\item Run \fn{/bin/rebaseall} inside the ash shell.
\end{enumerate}
Additional information on this problem can be found in the Cygwin
documentation.

The Cygwin versions of \Tcl/\Tk\ prior to 8.6 were not threaded, so
\OOMMF\ built with \Tcl/\Tk\ 8.5 and older will likewise not be
threaded.  This limitation is removed with the Cygwin \Tcl/\Tk\ 8.6
release.

\subsubsection{Setting the TCL\_LIBRARY environment variable}
\index{platform!Windows!setting~environment~variables}

If you encounter difficulties during \OOMMF\ start up, you may need to set
the environment variable
TCL\_LIBRARY\index{environment~variables!TCL\_LIBRARY}.  (NOTE: This
is almost never necessary!)

Bring up the Control Panel (e.g., by selecting
\btn{Settings\pipe Control Panel} off the Start menu), and select
\btn{System}.  Go to the \btn{Environment} tab, and enter
TCL\_LIBRARY as the Variable, and the name of the directory containing
\fn{init.tcl} for the Value, e.g.,
\begin{verbatim}
%SystemDrive%\Program Files\Tcl\lib\tcl8.0
\end{verbatim}
Click \btn{Set} and \btn{OK} to finish.

\blackhole{
\paragraph{On \Windows\ 9x}
Edit the file \fn{autoexec.bat}.  Add a line such as the following:
\begin{verbatim}
set TCL_LIBRARY=C:\Program Files\Tcl\lib\tcl8.0
\end{verbatim}
} % end-blackhole

%\subsubsection{Checking \fn{\bf .tcl} file association on \Windows~NT}
%\index{platform!Windows!file~extension~associations}
%As part of the \Tcl/\Tk\ installation, files with the \fn{.tcl}
%extension are normally associated with the \fn{wish} application.  This
%allows \Tcl\ scripts to be launched from
%\Windows\ Explorer\index{application!Windows~Explorer} by
%double-clicking on their icon, or from the NT command line without
%specifying the \fn{tclsh} or \fn{wish} shells.  If this is not working,
%you may check your installation from the NT command line as follows.
%First, run the command ``\cd{assoc~.tcl}''.  This should return the file
%type associated with the \fn{.tcl} extension, e.g., \cd{TclScript}.
%Next, use the \cd{ftype} command to check the command line associated
%with that file type, e.g.,
%%
%\begin{verbatim}
%C:\> ftype TclScript
% "C:\Program Files\Tcl\bin\wish84.exe" "%1" %2 %3 %4 %5 %6 %7 %8 %9
%\end{verbatim}
%%
%Note that the quotes are required as shown to protect spaces in
%pathnames.  If either \cd{assoc} or \cd{ftype} are incorrect, view the
%command line help information (``\cd{assoc~/?}'' and ``\cd{ftype~/?}'')
%for details on making changes.

%\subsubsection{Adding an \OOMMF\ shortcut to your desktop}
%\index{platform!Windows!desktop~shortcut}
%
%Right mouse click on the desktop to bring up the configuration dialog,
%and select \btn{New\pipe Shortcut}.
%Enter the command line necessary to bring up \OOMMF, e.g.,
%\begin{verbatim}
%tclsh84 c:\oommf\oommf.tcl
%\end{verbatim}

%Click \btn{Next\bca} and enter \cd{OOMMF} for the shortcut name.
%Select \btn{Finish}.

%At this point the shortcut will appear on your desktop with either the
%tclsh or wish icons.  Right mouse click on the icon and select
%\btn{Properties}.  Select the \btn{ShortCut} tab,
%and bring up \btn{Change Icon\ldots}  Under \btn{File Name:} enter the
%\OOMMF\ icon file, e.g.,
%\begin{verbatim}
%C:\oommf\oommf.ico
%\end{verbatim}
%
%Click \btn{OK}.  Back on the \btn{Shortcut} tab, change the
%\btn{Run:} selection to
%Minimized.  Click \btn{OK} to exit the Properties dialog box.  Double
%clicking on the \OOMMF\ icon should now bring up the
%\OOMMF\ application \app{mmLaunch}.
%\index{platform!Windows!configuration|)}

