\chapter{Data Table Display: mmDataTable}\label{sec:mmdatatable}%
\index{application!mmDataTable}

\begin{center}
\includepic{mmdatatable-ss}{mmDataTable Screen Shot}
\end{center}

\starsechead{Overview}
The application \app{mmDataTable} provides a data display service to its
client applications.  It accepts data from clients which are displayed in
a tabular format in a top-level window.  Its typical use is to display
the evolving values of quantities computed by micromagnetic solver
programs.

\starsechead{Launching}
{\bf mmDataTable} may be started either by selecting the
\btn{mmDataTable} button on \htmlonlyref{{\bf mmLaunch}}{sec:mmlaunch},
or from the command line via
\begin{verbatim}
tclsh oommf.tcl mmDataTable [standard options] [-net <0|1>]
\end{verbatim}

\begin{description}
\item[\optkey{-net \boa 0\pipe 1\bca}]
  Disable or enable a server which allows the data displayed by
  \app{mmDataTable} to be updated by another application.
  By default, the server is enabled.  When the server is disabled,
  \app{mmProbEd} is only useful if it is embedded in another application.
\end{description}

\starsechead{Inputs}
The client application(s) that send data to {\bf mmDataTable} for 
display control the flow of data.  The user, interacting with
the {\bf mmDataTable} window, controls how the data is displayed.
Upon launch, {\bf mmDataTable} displays only a menubar.  Upon user
request, a display window below the menubar displays data values.

Each message from a client contains a list of
(name, value, units) triples containing data for display.  
For example, one element in the list might be 
\cd{{\rm\{}Magnetization 800000 A/m{\rm\}}}.  {\bf mmDataTable}
stores the latest value it receives for each name.  Earlier
values are discarded when new data arrives from a client.

\starsechead{Outputs}
\app{mmDataTable} does not support any data output or storage
facilities.  To save tabular data, use the
\hyperrefhtml{\app{mmGraph}}{\app{mmGraph}
(Ch.~}{)}{sec:mmgraph}\index{application!mmGraph}
or
\hyperrefhtml{\app{mmArchive}}{\app{mmArchive}
(Ch.~}{)}{sec:mmarchive}\index{application!mmArchive} applications.

\starsechead{Controls}
The \btn{Data} menu holds a list of all the data names for which
\app{mmDataTable} has received data.  Initially, \app{mmDataTable} has
received no data from any clients, so this menu is empty.  As data
arrives from clients, the menu fills with the list of data names.
Each data name on the list lies next to a checkbutton.  When the
checkbutton is toggled from off to on, the corresponding data name and
its value and units are displayed at the bottom of the display window.
When the checkbutton is toggled from on to off, the corresponding data
name is removed from the display window.  In this way, the user
selects from all the data received what is to be displayed.  Selecting
the dashed rule at the top of the \btn{Data} menu detaches it so the
user may easily click multiple checkbuttons.

Displayed data values can be individually selected (or deselected) with
a left mouse button click on the display entry.  Highlighting is used to
indicated which data values are currently selected.  The {\btn{Options}}
menu also contains commands to select or deselect all displayed values.
The selected values can be copied into the
cut-and-paste\index{cut-and-paste} (clipboard) buffer with the
\key{CTRL-c} key combination, or the {\btn{Options\pipe Copy}} menu
command.

The data value selection mechanism is also used for data value
formatting control.  The \btn{Options\pipe Format} menu command brings
up a \wndw{Format} dialog box to change the justification and format
specification string.  The latter is the conversion string passed to
the \Tcl\ \cd{format} command, which uses the \C\ \cd{printf} format
codes.  If the {\btn{Adjust:Selected}} radiobutton is active, then the
specification will be applied to only the currently selected
(highlighted) data values.  Alternately, if {\btn{Adjust:All}} is
active, then the specification will be applied to all data values,
and will additionally become the default specification.

A right mouse button click on a display entry will select that entry,
and bring up the \wndw{Format} dialog box with the justification and
format specifications of the selected entry.  These specifications, with
any revisions, may then be applied to all of the selected entries.

If a value cannot be displayed with the selected format specification
string, e.g., if a ``\%d'' integer format were applied to a string
containing a decimal point, then the value will be printed in red
in the form as received by \app{mmDataTable}, without any additional
formatting.

The menu selection \btn{File\pipe Reset} reinitializes the
\app{mmDataTable} application to its original state, clearing the
display and the \btn{Data} menu.  The reset operation is also
automatically invoked upon receipt of new data following a data set
close message from a solver application.  The menu selection
\btn{File\pipe Exit} terminates the application.  The menu \btn{Help}
provides the usual help facilities.
