\chapter{Data Graph Display: mmGraph}\label{sec:mmgraph}%
\index{application!mmGraph}

\begin{center}
\includepic{mmgraph-ss}{mmGraph Screen Shot}
\end{center}

\starsechead{Overview}
The application \app{mmGraph} provides a data display service similar to
that of \hyperrefhtml{\app{mmDataTable}}{\app{mmDataTable}
(Ch.~}{)}{sec:mmdatatable}\index{application!mmDataTable}.  The usual
data source is a running solver, but rather than the textual output
provided by \app{mmDataTable}, \app{mmGraph} produces 2D line plots.
\app{mmGraph} also stores the data it receives, so it can produce
multiple views of the data and can save the data to disk.  Postscript
output is also supported.

\starsechead{Launching}
\app{mmGraph} may be started either by selecting the {\btn{mmGraph}}
button on \htmlonlyref{\app{mmLaunch}}{sec:mmlaunch} or from the command
line via
\begin{verbatim}
tclsh oommf.tcl mmGraph [standard options] [-net <0|1>] [loadfile ...]
\end{verbatim}

\begin{description}
\item[\optkey{-net \boa 0\pipe 1\bca}]
  Disable or enable a server which allows the data displayed by
  \app{mmGraph} to be updated by another application.
  By default, the server is enabled.  When the server is disabled,
  \app{mmGraph} may only input data from a file.
\item[\optkey{loadfile \ldots}]
  Optional list of data (\ODT) files to preload.
\end{description}

\starsechead{Inputs}
Input to \app{mmGraph} may come from either a file in the
\hyperrefhtml{\ODT\ format}{\ODT\ format
(Ch.~}{)}{sec:odtformat}\index{file!data~table},
or when \cd{-net 1} (the default) is active, from a client application
(typically a running solver).  The
\btn{File\pipe Open\ldots} dialog box is used to select an input file.
Receipt of data from client applications is the same as for
\hyperrefhtml{\app{mmDataTable}}{\app{mmDataTable}
(Ch.~}{)}{sec:mmdatatable}\index{application!mmDataTable}.  In either
case, input data are appended to any previously held data.

When reading from a file, \app{mmGraph} will automatically
decompress\index{compressed~files}
data using the \hyperrefhtml{local customization}{local customization
(Sec.~}{)}{sec:install.custom}\index{file!configuration}
``Nb\_InputFilter decompress'' option to \cd{Oc\_Option}.  For details,
see the discussion on file translation in the Inputs section of the
\hyperrefhtml{\app{mmDisp} documentation}{\app{mmDisp}
documentation (Ch.~}{)}{sec:mmdisp}.

Curve breaks\index{curve~break} (i.e., separation of a curve into
disjoint segments) are recorded in the data storage buffer via {\em
curve break records.}  These records are generated whenever a new data
table is detected by \app{mmGraph}, or when requested by the user using
the \app{mmGraph} \btn{Options\pipe Break Curves} menu option.

\starsechead{Outputs}

Unlike \app{mmDataTable}\index{application!mmDataTable}, \app{mmGraph}
internally stores the data sent to it.  These data may be written to
disk via the \btn{File\pipe Save As...} dialog box.  If the file
specified already exists, then {\bf mmGraph} output is appended to that
file.  The output is in the tabular \hyperrefhtml{\ODT\ format}{\ODT\
format described in Ch.~}{}{sec:odtformat}\index{file!data~table}.  The
data are segmented into separate \cd{Table Start/Table End} blocks
across each curve break record.

By default, all data currently held by \app{mmGraph} is written, but the
\btn{Save: Selected Data} option presented in the \btn{File\pipe Save
As...} dialog box causes the output to be restricted to those curves
currently selected for display.  In either case, the graph display
limits do {\em not} affect the output.

The save operation\index{data!save} writes records that are held by
\app{mmGraph} at the time the \btn{File\pipe Save As...} dialog box
\btn{OK} button is invoked.  Additionally, the \btn{Auto Save} option in
this dialog box may be used to automatically append to the specified
file each new data record as it is received by \app{mmGraph}.  The
appended fields will be those chosen at the time of the save operation,
i.e., subsequent changing of the curves selected for display does not
affect the automatic save operation.  The automatic save operation
continues until either a new output file is specified, the
\btn{Options\pipe Stop Autosave} control is invoked, or \app{mmGraph} is
terminated.

The \btn{File\pipe Print...}\index{data!print} dialog is used to
produce a Postscript file of the current graph.  On Unix systems, the
output may be sent directly to a
printer\index{platform!Unix!PostScript~to~printer}
by filling the \btn{Print to:}
entry with the appropriate pipe command, e.g., \texttt{|lpr}.  (The
exact form is system dependent.)

\starsechead{Controls}
Graphs are constructed by selecting any one item off the
\btn{X}-axis menu, and any number of items off the \btn{Y1}-axis and
\btn{Y2}-axis menus.  The y1-axis is marked on the left side of the
graph; the y2-axis on the right.  These menus may be detached by
selecting the dashed rule at the top of the list.  Sample results are
shown in the figure at the start of this section.

When \app{mmGraph} is first launched, all the axis menus are empty.
They are dynamically built based on the data received by \app{mmGraph}.
By default, the graph limits and labels are automatically set based on
the data.  The x-axis label is set using the selected item data label
and measurement unit (if any).  The y-axes labels are the measurement
unit of the first corresponding y-axis item selected.

The \btn{Options\pipe Configure...} dialog box allows the user to
override default settings.  To change the graph title, simply enter
the desired title into the \btn{Title} field.  To set the axis labels,
deselect the \btn{Auto Label} option in this dialog box, and fill in the
\btn{X Label}, \btn{Y1 Label} and \btn{Y2 Label} fields as desired.
The axis limits can be set in a similar fashion.  In addition, if an
axis limit is left empty, a default value (based on all selected data)
will be used.  Select the \btn{Auto Scale} option to have the axis
ranges automatically adjust to track incoming data.

Use the \btn{Auto Offset Y1} and \btn{Auto Offset Y2} to automatically
translate each curve plotted against the specified axis up or down so
that the first point on the curve has a y-value of zero for a linear
axis or one for a logarithmic axis.  This feature is especially useful
for comparing variations between different energy curves, because for
these curves one is typically interested in changes is values rather
than the absolute energy value itself.

By default the scaling on each axis is linear.  The \btn{Log scale axes}
check boxes enables logarithmic scaling on the selected axes.  If
logarithmic scaling is selected then any point with a zero value is
dropped and negative values are replaced with their corresponding
positive absolute value.

The size of the margin surrounding the plot region is computed
automatically.  Larger margins may be specified by filling in the
appropriate fields in the \btn{Margin Requests} section.  Units are
pixels.  Requested values smaller than the computed (default) values are
ignored.

The initial curve width is determined by the \cd{Ow\_GraphWin
  default\_curve\_width} setting in the \fn{config/options.tcl} and
\fn{config/local/options.tcl} files, following the usual method of
\hyperrefhtml{local customization}{local customization
  (Sec.~}{)}{sec:install.custom}\index{file!configuration}.  The current
curve width can be changed by specifying the desired width in the
\btn{Curve Width} entry in the {\btn{Options\pipe Configure...}} dialog
box.  The units are pixels.  Long curves will be rendered more quickly,
especially on \Windows, if the curve width is set to 1.

As mentioned earlier, \app{mmGraph} stores in memory all data it
receives.  Over the course of a long run, the amount of data stored can
grow to many megabytes\index{memory~use}.  This storage can be limited
by specifying a positive ($>0$) value for the \btn{Point buffer size}
entry in the {\btn{Options\pipe Configure...}} dialog box.  The oldest
records are removed as necessary to keep the total number of records
stored under the specified limit.  A zero value for \btn{Point buffer
size} is interpreted as no limit.  (The storage size of an individual
record depends upon several factors, including the number of items in
the record and the version of \Tcl\ being used.)  Data erasures may not
be immediately reflected in the graph display.  At any time, the point
buffer storage may be completely emptied with the \btn{Options\pipe
clear Data} command.  The \btn{Options\pipe Stop Autosave} selection
will turn off the auto save feature, if currently active.

Also on this menu is \btn{Options\pipe Rescale}, which autoscales the
graph axis limits from the selected data.  This command ignores but does
not reset the \btn{Auto Scale} setting in the \btn{Options\pipe
Configure...} dialog box.  The Rescale command may also be invoked by
pressing the \key{Home} key.

The \btn{Options\pipe Break Curves} item inserts a curve break record
into the point buffer, causing a break in each curve after the current
point.  This option may be useful if \app{mmGraph} is being fed data
from multiple sources.

The \btn{Options\pipe Key} selection toggles the key (legend) display on
and off.  The key may also be repositioned by dragging with the left
mouse button.  If curves are selected off both the y1 and y2 menus, then
a horizontal line in the key separates the two sets of curves, with the
labels for the y1 curves on top.

If the \btn{Options\pipe Auto Reset} selection is enabled, then when a
new table is detected all previously existing axis menu labels that are
not present in the column list of the new data set are deleted, along
with their associated data.  \app{mmGraph} will detect a new table when
results from a new problem are received, or when data is input from a
file.  If \btn{Options\pipe Auto Reset} is not selected, then no data or
axis menu labels are deleted, and the axes menus will show the union of
the old column label list and the new column label list.  If the axes
menus grow too long, the user may manually apply the
\btn{File\pipe Reset} command to clear them.

The last command on the options menu is \btn{Options\pipe Smooth}.  If
smoothing is disabled, then the data points are connected by straight
line segments.  If enabled, then each curve is rendered as a set of
parabolic splines, which do not in general pass through the data points.
This is implemented using the \cd{--smooth 1} option to the
\Tcl\ \cd{canvas create line} command; see that documentation for
details.

A few controls are available only using the mouse.  If the mouse pointer
is positioned over a drawn item in the graph, holding down the
\key{Control} key and the left mouse button will bring up the
coordinates of that point, with respect to the y1-axis.  Similarly,
depressing the \key{Control} key and the right mouse button, or
alternatively holding down the \key{Control}+\key{Shift} keys while
pressing the left mouse button will bring up the coordinates of the
point with respect to the y2-axis.  The coordinates displayed are the
coordinates of a point on a drawn line, which are not necessarily the
coordinates of a plotted data point.  (The data points are plotted at
the endpoints of each line segment.)  The coordinate display is cleared
when the mouse button is released while the \key{Control} key is down.

One vertical and one horizontal rule (line) are also available.
Initially, these rules are tucked and hidden against the left and bottom
graph axes, respectively.  Either may be repositioned by dragging with
the left or right mouse button.  The coordinates of the cursor are
displayed while dragging the rules.  The displayed y-coordinate
corresponds to the y1-axis if the left mouse button is used, or the
y2-axis if the right mouse button or the \key{Shift} key with the left
mouse button are engaged.

The graph extents may be changed by selecting a ``zoom box'' with the
mouse.  This is useful for examining a small portion of the graph in
more detail.  This feature is activated by clicking and dragging the
left or right mouse button.  A rectangle will be displayed that changes
size as the mouse is dragged.  If the left mouse button is depressed,
then the x-axis and y1-axis are rescaled to just match the extents of
the displayed rectangle.  If the right mouse button, or alternatively
the shift key + left mouse button, is used, then the x-axis and y2-axis
are rescaled.  An arrow is drawn against the rectangle indicating which
y-axis will be rescaled.  The rescaling may be canceled by positioning
the mouse pointer over the initial point before releasing the mouse
button.  The zoom box feature is similar to the mouse zoom control in
the \hyperrefhtml{\app{mmDisp}}{\app{mmDisp}
  (Ch.~}{)}{sec:mmdisp}\index{application!mmDisp} application, except
that here there is no ``un-zooming'' mouse control.  However, the
coordinate limits for each zoom are stored in a list, and the \key{Esc}
and \key{Shift+Esc} keys may be used to move backwards and forwards,
respectively, through this list.  The \key{Enter}/\key{Return} key will
copy the current coordinate limits to the list.  A pointer is kept to
the selected display configuration state, and new states are added after
that point; if a display state is added in the interior of the list
(i.e. ahead of the last state), then all configurations following the
new entry are deleted.  The entire list is cleared any time automatic
scaling is invoked.

The \key{PageUp} and \key{PageDown} keys may also be used to zoom the
display in and out.  Use in conjuction with the \key{Shift} key to jump
by larger steps, or with the \key{Control} key for finer control.  The
\btn{Options\pipe Rescale} command or the
\btn{Options\pipe Configure\ldots} dialog box may also be used to reset the
graph extents.

If \app{mmGraph} is being used to display data from a running solver,
and if \btn{Auto Scale} is selected in the
\btn{Options\pipe Configure\ldots} dialog box, then the graph extents
may be changed automatically when a new data point is received.  This is
inconvenient if one is simultaneously using the zoom feature to examine
some portion of the graph.  In this case, one might prefer to disable
the \btn{Auto Scale} feature, and manually pan the display using the
keyboard arrow keys.  Each key press will translate the display one half
frame in the indicated direction.  The \key{Shift} key used in
combination with an arrow keys double the pan step size, while the
\key{Control} key halves it.

The menu selection \btn{File\pipe Reset} reinitializes the
\app{mmGraph} application to its original state, releasing all data and
clearing the axis menus.  The menu selection \btn{File\pipe Exit}
terminates the application.  The menu \btn{Help} provides the usual help
facilities.

\starsechead{Details}
The axes menus are configured based on incoming data.  As a result,
these menus are initially empty.  If a graph widget is scheduled to
receive data only upon control point or stage done events in the solver,
it may be a long time after starting a problem in the solver before the
graph widget can be configured.  Because \app{mmGraph} keeps all data up
to the limit imposed by the \cd{Point buffer size}, data loss is usually
not a problem.  Of more importance is the fact that automatic data
saving\index{data!save} can not be set up until the first data point is
received.  As a workaround, the solver initial state may be sent
interactively as a dummy point to initialize the graph widget axes
menus.  Select the desired quantities off the axes menus, and use the
\btn{Options\pipe clear Data} command to remove the dummy point from
\app{mmGraph}'s memory.  The {\btn{File\pipe Save As...}} dialog box may
then be used---with the {\btn{Auto Save}} option enabled---to write out
an empty table with proper column header information.  Subsequent data
will be written to this file as they arrive.


