\chapter{Documentation Viewer: mmHelp}\label{sec:mmhelp}
\index{application!mmHelp}

\begin{center}
\includepic{mmhelp-ss}{mmHelp Screen Shot}
\end{center}

\starsechead{Overview}
The application {\bf mmHelp} manages the display and navigation of
hypertext (HTML) help files\index{file!HTML}.  
%It presents an interface similar
%to that of World Wide Web browsers\index{application!web~browser}.
%
%Although {\bf mmHelp} is patterned after World
%Wide Web browsers, it does not have all of their capabilities.
{\bf mmHelp} displays only a simplified form of hypertext
required to display the \OOMMF\ help pages.
It is not able to display many of
the advanced features provided by modern World Wide Web
browsers.  
%In the current release, {\bf mmHelp} is not able to 
%follow {\tt http:} URLs.  It only follows {\tt file:} URLs\index{URL},
%such as
%\begin{center}
%file:/path/to/oommf/doc/userguide/userguide/Documentation\_Viewer\_mmHelp.html
%\end{center}
\OOMMF\ software can be \hyperrefhtml{customized}{customized (See
Sec.~}{)}{sec:install.custom} to use another program to display the HTML
help files\index{customize!help~file~browser}.

\starsechead{Launching}
{\bf mmHelp} may be launched from the command line via
\begin{verbatim}
tclsh oommf.tcl mmHelp [standard options] [URL]
\end{verbatim}
The command line argument {\tt URL} is the URL of the first
page (home page) to be displayed.  If no URL is specified,
{\bf mmHelp} displays the Table of Contents of the {\em \OOMMF\ User's
Guide} by default.

\starsechead{Controls}
Each page of hypertext is displayed in the main {\bf mmHelp} window.
Words which are underlined and colored blue are hyperlinks which {\bf
mmHelp} knows how to follow.  Words which are underlined and colored red
are hyperlinks which {\bf mmHelp} does not know how to follow.  Moving
the mouse over a hyperlink displays the target URL of the hyperlink in
the \btn{Link:} line above the display window.  Clicking on a blue
hyperlink will follow the hyperlink and display a new page of hypertext.

{\bf mmHelp} keeps a list of the viewed pages in order of view.
Using the \btn{Back} and \btn{Forward} buttons, 
the user may move backward and forward through 
this list of pages.
The \btn{Home} button causes the first page to be displayed, 
allowing the user to start again from the beginning.  These
three buttons have corresponding entries in the 
\btn{Navigate} menu.

Use the menu selection \btn{File\pipe Open} to directly select
a file from the file system to be displayed by {\bf mmHelp}.

The menu selection \btn{File\pipe Refresh}, or 
the \btn{Refresh} button causes {\bf mmHelp} to reload and
redisplay the current
page.  This may be useful if the display becomes corrupted,
or for repeatedly loading a hypertext file which is being
edited.

When {\bf mmHelp} encounters hypertext elements it does not
recognize, it will attempt to work around the problem.  
However, in some cases it will not be able to make sense of 
the hypertext, and will display an error message.  Documentation
authors should take care to use only the hypertext elements
supported by {\bf mmHelp} in their documentation files.  Users
should never see such an error message.

{\bf mmHelp} displays error messages in one of two ways: 
within the display window, or in a separate window.  
Errors reported in the display window replace the display 
of the page of hypertext.
They usually indicate that the hypertext page could not be
retrieved, or that its contents are not hypertext.  File
permission errors are also reported in this way.

Errors reported in a separate window are usually due to a
formatting error within the page of hypertext.  Selecting the 
\btn{Continue} button of the error window instructs {\bf mmHelp} to 
attempt to resume display of the hypertext page beyond the error.  
Selecting \btn{Abort} abandons further display.  

The menu selection \btn{Options\pipe Font scale...} brings up a
dialog box through which the user may select the scale of
the fonts to use in the display window, relative to their
initial size.

\blackhole{
The menu selection \btn{Options\pipe Strict} should not
normally be enabled.  Enabling it causes {\bf mmHelp} to
become much more finicky about the correctness of the HTML
in the pages it displays.  The main consequence of enabling
this option is that {\bf mmHelp} will display more error
messages.  This option is primarily useful when \OOMMF\
documentation writers are writing HTML files directly,
and wish to check their work.
} % end-blackhole

The menu selection \btn{File\pipe Exit} or the \btn{Exit} button
terminates the {\bf mmHelp} application.  
The menu \btn{Help} provides the usual help facilities.

\starsechead{Known Bugs}
\app{mmHelp} is pretty slow.  You may be happier using
\hyperrefhtml{local customization}{local
customization (Sec.~}{)}{sec:install.custom} methods to replace it
with another more powerful HTML browser.  Also, we have noticed that the
underscore character in the italic font is not displayed (or is
displayed as a space) at some font sizes on some platforms.

