\section*{Disclaimer}
\addcontentsline{toc}{section}{Disclaimer}

The research software described in this manual (``software'') is provided
by NIST as a public service. You may use, copy and distribute copies of
the software in any medium, provided that you keep intact this entire notice.
You may improve, modify and create derivative works of the software or any
portion of the software, and you may copy and distribute such modifications
or works. Modified works should carry a notice stating that you changed the
software and should note the date and nature of any such change. Please
explicitly acknowledge the National Institute of Standards and Technology
as the source of the software.

The software is expressly provided "AS IS." NIST MAKES NO WARRANTY OF ANY
KIND, EXPRESS, IMPLIED, IN FACT OR ARISING BY OPERATION OF LAW, INCLUDING,
WITHOUT LIMITATION, THE IMPLIED WARRANTY OF MERCHANTABILITY, FITNESS FOR A
PARTICULAR PURPOSE, NON-INFRINGEMENT AND DATA ACCURACY. NIST NEITHER
REPRESENTS NOR WARRANTS THAT THE OPERATION OF THE SOFTWARE WILL BE
UNINTERRUPTED OR ERROR-FREE, OR THAT ANY DEFECTS WILL BE CORRECTED. NIST
DOES NOT WARRANT OR MAKE ANY REPRESENTATIONS REGARDING THE USE OF THE
SOFTWARE OR THE RESULTS THEREOF, INCLUDING BUT NOT LIMITED TO THE
CORRECTNESS, ACCURACY, RELIABILITY, OR USEFULNESS OF THE SOFTWARE. 

You are solely responsible for determining the appropriateness of using
and distributing the software and you assume all risks associated with
its use, including but not limited to the risks and costs of program errors,
compliance with applicable laws, damage to or loss of data, programs or
equipment, and the unavailability or interruption of operation. This
software is not intended to be used in any situation where a failure
could cause risk of injury or damage to property.  The software was
developed by NIST employees. NIST employee contributions are not subject
to copyright protection within the United States. 

We would appreciate acknowledgement if the software is used.  When
referencing \OOMMF\ software, we recommend citing the NIST technical
report, M. J. Donahue and D. G. Porter, ``OOMMF User's Guide, Version
1.0,'' \textbf{NISTIR 6376}, National Institute of Standards and
Technology, Gaithersburg, MD (Sept 1999).

Commercial equipment and software referred to on these pages are
identified for informational purposes only, and does not imply
recommendation of or endorsement by the National Institute of Standards
and Technology, nor does it imply that the products so identified are
necessarily the best available for the purpose.
